---
title: "How you Lose your Money in the Roulette"
output: html_notebook
---

<!-- code_folding is for present with hiding code in markdown code_folding: hide-->

```{r cache=TRUE,message=FALSE, warning=FALSE}
library(dplyr)
#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list
one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 


# Simulation
walk_out_money <- rep(NA, 1000)
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_last
}

# Walk out money distribution
hist(walk_out_money, breaks = 100)

# Estimated probability of walking out with extra cash
mean(walk_out_money > 200)

# Estimated earnings
mean(walk_out_money - 200)

```
  
   

ANSWER
### You should explain how you used computer simulation to calculate the average earnings of a gambler that uses this strategy. As part of the explanation, provide a figure (or a series of figures) that show how the gamblers earnings (or losses) evolve over a series of wagers at the roulette wheel. (The x-axis will be the wager number (or play number), the y-axis will be earnings.) The code below provides all the functions you’ll need to calculate average earnings.

```{r cache=TRUE,message=FALSE, warning=FALSE}
earnings <- one_series(B = 200, W = 300, L = 1000, M = 100) - 200

plot(earnings)
```



### Show your audience how changing a parameter of the simulation (see table below) does or does not have an impact on average earnings. A figure would be helpful.

```{r cache=TRUE,message=FALSE, warning=FALSE}
B <- seq(50,500,50)
avg_earnings <- rep(NA, length(B))
for(i in 1:length(B)){
  walk_out_money <- rep(NA, 1000)
  for(j in seq_along(walk_out_money)){
    walk_out_money[j] <- one_series(B = B[i], W = 300, L = 1000, M = 100) %>% get_last
  }
  avg_earnings[i] <- (mean(walk_out_money)-B)
}

plot(B, avg_earnings)
```

### See the stopping rule below. Explain to your audience how you used computer simulation to estimate the average number of plays before stopping. The code below will need to be modified to calculate this quantity.

If the player wins more than $100 then it will stop. If the player loses all the money he brings to the table then it will stop. If the player has been spinned for more than 1000 times,it will stop as well. 

```{r cache=TRUE,message=FALSE, warning=FALSE}
play_times <- rep(NA, 1000)
for(j in seq_along(play_times)){
  play_times[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% length
}
```


### Be sure to explain the limitations of the simulation; identify simplifications or other sources of uncertainty.
It only runs 1000 times. This simulation doesn't count the back box cases in. In the reality, things may be more comlicated and unpredictable.
