---
title: "method-comparison"
date: "10/28/2019"
author: "Wenying Gu"
output:
  html_document:
    code_folding: hide
    theme: united 
    highlight: tango
    toc: yes
    toc_depth: 6 
    toc_float: true
editor_options: 
  chunk_output_type: inline
---
### Email: [wenying.gu@vanderbilt.edu]()

# Simulation study
Coverage probability is an important operating characteristic of methods for constructing interval estimates, particularly confidence intervals. (See the previous deliverable for definitions and examples.)

Idealy, a 95% confidence interval will capture the population parameter of interest in 95% of samples. One can also calculate 80% or 90% confidence intervals. In general, an X% confidence interval should capture the population parameter of interest in X% of samples.

# Aim for this blog 
This blog will perform a 2 × 4 × 2 factorial simulation study to compare the coverage probability of various methods of calculating 90% confidence intervals. The three factors in the experiment are

1.True, underlying distribution

* standard normal
* gamma(shape = 1.4, scale = 3)

2.Model

* method of moments with normal
* method of moments with gamma
* kernel density estimation
* bootstrap

3.Parameter of interest

* sample min (1st order statistic)
* median

Other settings in the experiment that will not change are:

* Sample size, N = 201
* Outside the loop estimation

# Library 
```{r cache=TRUE,message=FALSE, warning=FALSE}
library(dplyr)
require(foreach)
require(doParallel)
```

# Generate data 
Write a function to generate data from standard normal distribution and gamma distribution of size N = 201. Generating the data use rnorm(N) and rgamma(N,shape,scale).

1.True, underlying distribution

* standard normal
* gamma(shape = 1.4, scale = 3)

```{r cache=TRUE,message=FALSE, warning=FALSE}

generate_data <- function(N, dist, sh, sc) {
  if (dist == "norm") {
    rnorm(N)
  } else if (dist == "gamma") {
    rgamma(N, shape = sh, scale = sc)
  }
}

```

# Use Different Methods to Estimate the distribution 
Build a model to estimate 90% confidence interval for the distribution min/median, the model will contain 4 methods, including method of moments with normal; method of moments with gamma; kernel density estimation and bootstrap. 

2.Model

* method of moments with normal
* method of moments with gamma
* kernel density estimation
* bootstrap

3.Parameter of interest

* sample min (1st order statistic)
* median

```{r cache=TRUE,message=FALSE, warning=FALSE}

estimate.ci <- function(data, mod, par.int, R = 500, smoo = 0.3) {
  N<- length(data)
  sum.measure <- get(par.int1)

  
  if (mod == "MMnorm") {
    mm.mean <- mean(data)
    mm.sd <- sd(data)
    sim.data <- array(rnorm(length(data)*R), dim = c(N, R))
    samp.dist <- apply(sim.data, 2, FUN = sum.measure)
    return(quantile(samp.dist, c(0.05, 0.95)))
    
  } else if (mod == "MMgamma") {
    
    mm.shape <- mean(data)^2 / var(data)
    mm.scale <- var(data) / mean(data)
    sim.data <- array(rgamma(length(data)*R, shape = mm.shape, scale = mm.scale), dim = c(N, R))
    samp.dist <- apply(sim.data, 2, FUN = sum.measure)
    return(quantile(samp.dist, c(0.05, 0.95)))
    
  } else if (mod == "KDE") {
    
    ecdfstar <- function(t, data, smooth = smoo) {
      outer(t, data, function(a,b){pnorm(a, b, smooth)}) %>% rowMeans
    }
    
    tbl <- data.frame(
      x = seq(min(data)-sd(data), max(data)+sd(data), by = 0.01)
    ) 
    
    tbl$p <- ecdfstar(tbl$x, data, smoo) 
    tbl <- tbl[!duplicated(tbl$p), ]
    
    qkde <- function(ps, tbl){
      rows <- cut(ps, tbl$p, labels = FALSE) 
      tbl[rows, "x"]
    }
    
    U <- runif(N*R)
    sim.data <- array(qkde(U, tbl), dim = c(N, R))
    samp.dist <- apply(sim.data, 2, FUN = sum.measure)
    return(quantile(samp.dist, c(0.05, 0.95), na.rm = TRUE))
    
  } else if (mod == "Boot") {
  
    sim.data <- array(sample(data, N*R, replace = TRUE), dim = c(N, R))
    samp.dist <- apply(sim.data, 2, FUN = sum.measure)
    return(quantile(samp.dist, c(0.05, 0.95)))
  }
}


```

# Capture the min/median 
If the sample size is large enough, our min/median will close to the theoretical value. Therefore we want the know whether confidence interval capture the theoretical value(true min/median). 

Capture parameter: the lower confidence limit is less than true parameter and the upper confidence limit is greater than true parameter.

```{r cache=TRUE,message=FALSE, warning=FALSE}
capture_par <- function(ci, true.par) {
  1 * (ci[1] < true.par & true.par < ci[2])
}
```


# Estimate the coverage probability 
Coverage probability refers to the probability that capturing the true min/median. In the simulation, it will return 1 if the confidence interval capture the true min/median else will return 0. Therefore, take the mean of the returns can get the coverage probability. 

```{r cache=TRUE,message=FALSE, warning=FALSE}
N <- 201
shape.set <- 1.4
scale.set <- 3

true.norm.med <- qnorm(0.5)
true.norm.min <- mean(apply(array(rnorm(N*10000), dim = c(N, 10000)), 2, min))
true.gamma.med <- qgamma(0.5, shape = shape.set, scale = scale.set)
true.gamma.min <- mean(apply(array(rgamma(N*10000, shape = shape.set, scale = scale.set), dim = c(N, 10000)), 2, min))



simsettings <- expand.grid(dist = c("norm", "gamma"), model = c("MMnorm", "MMgamma", "KDE", "Boot"), par.int = c("median", "min"), cov.prob = NA, stringsAsFactors = FALSE, KEEP.OUT.ATTRS = FALSE)

```

```{r message=FALSE}
library(dplyr)

for (k in c(1:2, 4:10,12:16)) {
  dist1 <- simsettings[k, 1]
  model1 <- simsettings[k, 2]
  par.int1 <- simsettings[k, 3]
  
  if(dist1=="norm" & par.int1=="median"){
    true.par1 = true.norm.med
  }else if(dist1=="norm" & par.int1=="min"){
    true.par1 = true.norm.min
  }else if(dist1=="gamma" & par.int1=="median"){
    true.par1 = true.gamma.med
  }else if(dist1=="gamma" & par.int1=="min"){
    true.par1 = true.gamma.min
  }

cover <- NA
for (sims in 1:500) {
  cover[sims] <- generate_data (N, dist1,1.4,3)%>% estimate.ci(mod = model1, par.int = par.int1, R=500,smoo = 0.3) %>% capture_par(true.par = true.par1)
}
simsettings[k,4] <- mean(cover)
}
library(kableExtra)
kable(simsettings)%>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

# Conclusion
The combination model doesn't work since the gamma distribution can not contain negative scale. Here is the shape and scale value under MMgamma function.

*Browse[1]> mm.shape
*[1] 2.914494e-07
*Browse[1]> mm.scale
*[1] -1895.302

We can add the value to "shape" to make it positive and the function will work. 