---
title: "08-coverage-probability"
date: "10/15/2019"
author: "Wenying Gu"
output:
  html_document:
    code_folding: hide
    theme: united 
    highlight: tango
    toc: yes
    toc_depth: 6 
    toc_float: true
editor_options: 
  chunk_output_type: inline
---
### Email: [wenying.gu@vanderbilt.edu]()

# What is coverage probability?
Coverage probability is an important operating characteristic of methods for constructing interval estimates, particularly confidence intervals.

# Aims for This Blog 
Generate a simulation to calculate the coverage probability of the 95% confidence interval of the median by using MLE method. 

# Generate a single sample data 
Generate a single sample from a standard normal distribution of size N = 201. Generating the data use rnorm and then construct the confidence interval. 
```{r}
N<-201
require(stats4)
parameters<-c(N=201, mean=0, sd=1)
generate_data<-rnorm(201)
construct_ci <- function(mean, sd){
  fs <- dnorm(
        x = generate_data
      , mean = mean
      , sd = sd
      , log = TRUE
    ) 
  -sum(fs,na.rm = TRUE)
}

```

# Use MLE to estimate the distribution 
Estimating the sampling distribution of the single sample by using the MLE method from package stats4. Inputing the mean =0 and sd=1 to find the estimatie mean and sd by MLE method. Usually, the mean is around 0 and the sd is around 1. 

```{r}

fit <- mle(
    construct_ci
  , start = list(mean = 0, sd = 1)
  , method = "L-BFGS-B"
  , lower = c(0, 0.01)
)

capture_param<-coef(fit)
capture_param

```

# Get confidence interval for the median 
Set up a simulation to run the previous steps for 5000 times and get 5000 median. Using the quantile() to find the 95% confidence interval which is the true mean of 5000 median samples.  
```{r}
median <- rep(NA, 5000)
for(i in 1:5000){
  b <- rnorm(201, mean = stats4::coef(fit)[1], sd =stats4::coef(fit)[2] ) 
  median[i]<-median(b)
}

```


```{r}
quantile95 <- quantile(median, c(0.025,0.975))
quantile95
```

# Capture the median 
The sampling distribution of the median is likely to be normally distributed when the sample size 'n' is large. The median for standard normal distribution is 0. Therefore we want the know whether confidence interval capture the true median. 

True median: the lower confidence limit is less than zero and the upper confidence limit is greater than zero.

```{r}
N <- 201 
pop.mean = 0
pop.sd= 1


pop.sample<-rnorm(N,0,1)
mle.sd <- sqrt(((N-1)/N)*var(pop.sample))
mle.mean <- mean(pop.sample)

capture <- NA
for(j in 1:1000) {
  out <- rep(NA, 1000)
  for(i in 1:1000) {
    Fhat.sample <- rnorm(N, mean = mle.mean, sd = mle.sd)
    out[i] <- median(Fhat.sample)
  }
  gen.ci <- quantile(out, c(0.025, 0.975))
  capture[j] <-(gen.ci[1] < pop.mean & gen.ci[2] > pop.mean)
}

```

#Estimate the coverage probability 
Coverage probability refers to the probability that capturing the true median. In the simulation, it will return 1 if the confidence interval capture the true median else will return 0. Therefore, take the mean of the returns can get the coverage probability. MLE is a good way to estimate the normal distribution, our covergae probability is around 1. 
```{r}
round(mean(capture), 4)
```

